### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9b5e8dc75883cba38d2bbae428327ca82b972ccd 🚀

Spillover risk is multidimensional and complicated. From the standpoint
of an animal community, \emph{i.e.} a pool of suitable hosts, it is
driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
the global richness of hosts is one such component commonly analysed
(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
argument to be made that species who are not competent (or known) hosts
of a specific virus genus may not factor into this (Plowright et al.
2015); this is especially true as competence data increases predictive
power when the taxonomic scope of hosts of a viral family increases
(Becker et al. 2020, Mull et al. 2022). Similarly, host species who are
assumed to share viruses at different rates should be weighted
accordingly (Albery et al. 2020). In mammals, key functional traits (for
which phylogeny is a reasonable proxy) are determinants of the spillover
potential (Olival et al. 2017); these include, notably, body mass and
affinity for urban environments (Albery et al. 2022). Finally,
especially when the pool of potential hosts spans the entire globe,
there may be local host pools that are highly unique; not having been
observed in other locations, these can act on the overall risk either by
providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens (which is the
case many viruses in the betacoronavirus genus, see \emph{e.g.} MacLean
et al. 2021), there is conceptual and empirical support to the idea that
these community-level mechanisms are even more important in driving the
overall risk (Power and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
threaten human health (Letko et al. 2020, Van Brussel and Holmes 2022).
Chiropterans emerged around 64 million years ago and are one of the most
diverse mammalian orders, with an estimated richness of more than 1400
species (Peixoto et al. 2018, Simmons and Cirranello 2020). They exhibit
a broad variety of habitat use, behaviour, and feeding strategies,
putting them at key positions in the delivery and provisioning of
several ecosystem services, tied to important ecosystem-derived benefits
to human (Kasso and Balakrishnan 2013). For example, bats are an
essential component of many seed-dispersal networks (Mello et al. 2011).
Over two-thirds of bats are know to be either obligate or facultative
insectivores, therefore playing an important role in the regulation of
insect pests that can affect crops (Williams-Guillén et al. 2008, Voigt
and Kingston 2016), and vectors of pathogens that put a risk on human
health (Gonsalves et al. 2013a, b). Because bats are globally
distributed and have a long evolutionary history, phylogeographic and
biogeographic approaches are required to shed light on the contemporary
distribution of coevolutionary processes between bats and the pathogens
they host. Not all areas in which bats, viruses, and human are
co-occuring are facing a risk of spillover towards human populations,
and the areas in which this risk exist may not be facing risks of the
same nature and magnitude.

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of all
confirmed bat hosts of betacoronaviruses. This work is important both as
a description of the bats-betacoronavirus complex, but aslo because more
broadly, bats are known reservoirs for a variety of emerging viruses and
other pathogens (Calisher et al. 2006, Melaun et al. 2014), making
balancing the needs for bat conservation and disease prevention most
likely very difficult and a source of human-wildlife conflicts,
especially in more densely populated areas (Stone et al. 2015, Rego et
al. 2015). By drawing on concepts from the Geographic Mosaic Theory of
Coevolution (Thompson 2005), we turn these associations into a spatially
explicit additive mapping of zoonotic risk components, revealing the
extreme heterogeneity of risk at the global scale. We identify the
Amazon and South-Eastern Asia as hotspots where the phylogenetic
distinctiveness of betacoronaviruses is the highest (Anthony et al.
2017); surprisingly, current data suggest that viral sharing between
hosts is high in the Amazon and low in South-Eastern Asia, which has the
potential to result in different evolutionary dynamics between these two
regions.

@@

>>>>

Spillover risk is multidimensional and complicated. Within a pool of
competent hosts, it is driven by a multiplicity of factors (Plowright et
al. 2017). Although proxies for the local richness of hosts is commonly
analysed (see \emph{e.g.} Anthony et al. 2017 for coronaviruses), there
is an argument to be made that species who are not (known as) competent
hosts of a specific virus genus may not factor into this (Plowright et
al. 2015), calling for species-level information. This is especially
true as competence data increases predictive power when the taxonomic
scope of hosts of a viral family increases (Becker et al. 2020, Mull et
al. 2022). Similarly, host species who share viruses at different rates
should be weighted accordingly (Albery et al. 2020). In mammals, key
functional traits (for which phylogeny is a reasonable proxy) are
determinants of the spillover potential (Olival et al. 2017); these
include, notably, body mass, and affinity for urban environments (Albery
et al. 2022). Finally, especially when the pool of potential hosts spans
the entire globe, there may be local host pools that are highly unique;
not having been observed in other locations, these can act on the
overall risk either by providing novel contact opportunities, reflecting
unique host-environment combinations (Engering et al. 2013), or
facilitating rapid evolutionary changes in specialism of their pathogens
(Agosta et al. 2010). In the specific case of generalist pathogens
(which is the case many viruses in the betacoronavirus genus, see
\emph{e.g.} MacLean et al. 2021), there is conceptual and empirical
support to the idea that these community-level mechanisms are even more
important in driving the overall risk (Power and Mitchell 2004).

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of all
confirmed bat hosts of betacoronaviruses. By drawing on concepts from
the Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005), we
turn these associations into a spatially explicit additive mapping of
zoonotic risk components, revealing the extreme heterogeneity of risk at
the global scale. Explicitely framing the notion of spillover risk based
on propositions from the GMTC (which is to say, based on a framework
linking interactions between species to change within species) is a
novel idea, that should be relatively general. Indeed, it only assumes
the action of well described evolutionary mechanisms. The benefit of
this approach is to provide the potential for a more dynamic and nuanced
understanding of risk: not only on ecological timescales, but also by
providing clues about which areas can change over micro-evolutionary
timescales. This provides a way to look at spatial structure by
accounting for more notions than species richness/similarity, but also a
way to identify spatial areas of higher risk.

We identify the Amazon and South-Eastern Asia as hotspots where the
phylogenetic distinctiveness of betacoronaviruses is the highest
(Anthony et al. 2017); surprisingly, current data suggest that viral
sharing between hosts is high in the Amazon and low in South-Eastern
Asia, which has the potential to result in different evolutionary
dynamics between these two regions, hinting at different futures for
their viral communities. This work is important both as a description of
the bats-betacoronavirus complex, but also because more broadly, bats
are known reservoirs for a variety of emerging viruses and other
pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
the needs for bat conservation and disease prevention most likely very
difficult and a source of human-wildlife conflicts, especially in more
densely populated areas (Stone et al. 2015, Rego et al. 2015).

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human health (Letko et al.
2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
million years ago and are one of the most diverse mammalian orders, with
an estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
know to be either obligate or facultative insectivores, therefore
playing an important role in the regulation of insect pests that can
affect crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016),
and vectors of pathogens that put a risk on human health (Gonsalves et
al. 2013a, b). Because bats are globally distributed and have a long
evolutionary history, phylogeographic and biogeographic approaches are
required to shed light on the contemporary distribution of
coevolutionary processes between bats and the pathogens they host. Not
all areas in which bats, viruses, and human are co-occuring are facing a
risk of spillover towards human populations, and the areas in which this
risk exist may not be facing risks of the same nature and magnitude.
